RSV

### 2018-01-02 Analysis/RSV.Rmd
Refactoring code for paper into separate R file

library(reshape)
library(plyr)
library(dplyr)
library(zipcode)
library(mgcv)
library(tsgam)
library(ggplot2)
library(easyGgplot2)
library(utils)
library(outbreakpredict)
data(zipcode)
set.seed(0)
```

```{r}
rsvWeek = function(weeki) {
  (weeki + 26) %% 52 - 26
}

dataset = read.csv("../ct rsv time series.csv", header=TRUE, colClasses=c("patzip"="character"))
# Add week-of-year, with zero = end of June, drop first 5.5 years because of coding differences
dataset = dataset %>%
  mutate(weekiadj=round(dataset$weeki * 364 / 365.25), weekofyear=rsvWeek(weekiadj), year=1990 + (weekiadj + 26) %/% 52) %>%
  filter(weeki > 5.5*52) %>%
  filter(weeki < 22.5*52)
# Add state based on zipcode and drop everything not in CT
zipState = zipcode %>% select(zip, state)
dataset = dataset %>% 
  left_join(zipcode %>% select(zip, state), by=c("patzip"="zip")) %>% 
  filter(state=="CT")
# Add county based on zipcode
countyByZip = read.csv("../ct counties.csv", header=TRUE, colClasses=c("zip"="character"))
counties = countyByZip %>% select(county) %>% distinct()
dataset = dataset %>% inner_join(countyByZip, by=c("patzip"="zip"))

# Sort counties by total RSV
totalRSVByCounty = dataset %>%
  group_by(county) %>% 
  summarize(total_rsv=sum(rsv)) %>%
  arrange(total_rsv) %>%
  mutate(county=factor(county, county))

rsvCounties = totalRSVByCounty$county
rsvYears = sort(unique(dataset$year))
rsvWeekRange = sort(unique(dataset$weeki))
rsvWeekOfYearRange = sort(unique(dataset$weekofyear))

dataset = dataset %>%
  mutate(county=factor(county, levels=levels(rsvCounties)))

@@

```{r include=FALSE}
# Reshape counties into columns and add column for total
rsvByCountyAndWeekOfYear = dataset %>%
  group_by(county, weekofyear) %>%
  summarize(rsv=sum(rsv)) %>%
  group_by(weekofyear) %>%
  do(rbind(., as.data.frame(.) %>%
             summarize(county="total", weekofyear=max(weekofyear), rsv=sum(rsv))
           ))

rsvForAllCountiesByWeekOfYear = rsvByCountyAndWeekOfYear %>%
  melt(id=c("weekofyear", "county"), measured=c("rsv")) %>%
  cast(weekofyear ~ county + variable) %>%
  rename_all(funs(gsub("_rsv", "", .)))

maxRSVByCounty = rsvByCountyAndWeekOfYear %>%
  group_by(county) %>% 
  summarize(max_rsv=max(rsv))

rsvRelForAllCountiesByWeekOfYear = rsvByCountyAndWeekOfYear %>%
  inner_join(maxRSVByCounty, by="county") %>%
  mutate(rsv_rel=rsv/max_rsv) %>%
  select(-max_rsv, -rsv) %>%
  melt(id=c("weekofyear", "county"), measured=c("rsv_rel")) %>%
  cast(weekofyear ~ county + variable)

allCountiesTS = rsvForAllCountiesByWeekOfYear %>%
  arrange(weekofyear) %>%
  select(-weekofyear) %>%
  ts(frequency=1, start=c(-26, 1))

allCountiesRelTS = rsvRelForAllCountiesByWeekOfYear %>%
  arrange(weekofyear) %>%
  select(-weekofyear) %>%
  ts(frequency=1, start=c(-26, 1))

plot.ts(allCountiesTS, nc=1)
plot.ts(allCountiesRelTS, nc=1, plot.type="single", col=rainbow(8), lwd=2)
```

```{r include=FALSE}
# Drop counties in which max # of RSV cases in a week is <100
maxRSVByInterestingCounty = maxRSVByCounty %>% filter(max_rsv > 50)

rsvForInterestingCountiesByWeekOfYear = rsvByCountyAndWeekOfYear %>%
  inner_join(maxRSVByInterestingCounty, by="county") %>%
  select(-max_rsv) %>%
  melt(id=c("weekofyear", "county"), measured=c("rsv", "rsv_rel")) %>%
  cast(weekofyear ~ county + variable)

rsvRelForInterestingCountiesByWeekOfYear = rsvByCountyAndWeekOfYear %>%
  inner_join(maxRSVByInterestingCounty, by="county") %>%
  mutate(rsv_rel=rsv/max_rsv) %>%
  select(-max_rsv, -rsv) %>%
  melt(id=c("weekofyear", "county"), measured=c("rsv_rel")) %>%
  cast(weekofyear ~ county + variable)

interestingCountiesTS = rsvForInterestingCountiesByWeekOfYear %>%
  arrange(weekofyear) %>%
  select(-weekofyear) %>%
  ts(frequency=1, start=c(-26, 1))

interestingCountiesRelTS = rsvRelForInterestingCountiesByWeekOfYear %>%
  arrange(weekofyear) %>%
  select(-weekofyear) %>%
  ts(frequency=1, start=c(-26, 1))

plot.ts(interestingCountiesTS, nc=1)
plot.ts(interestingCountiesRelTS, nc=1, plot.type="single", col=rainbow(5), lwd=2)
```

```{r}
outbreak.calc.unprotected = function(model, params, time, strategies) {
  # Get model predictions for given (randomized) param values
  time = time + 0.5
  predictors = model %>% predict(data.frame(time=time), type="lpmatrix")
  fit = predictors %*% params

  # Map spline fit back to data
  fit = fit %>% model$family$linkinv()

  # Calculate total # of unprotected cases for each strategy
  unprotected = as.data.frame(lapply(strategies, function(strat) {
    sum(fit * (1 - sapply(time, strat)))
  }))

  # Calculate total # of cases 
  total = sum(fit)

  unprotected %>% 
    rename_all(funs(
      sprintf("%s.count", .)
    )) %>%
    cbind(
      (unprotected / total) %>%
        rename_all(funs(
          sprintf("%s.frac", .)
        ))
    ) %>% 
    mutate(total=total)
}
```

```{r}
eps = .05
seasonThreshold = 0.025
rsvTime = dataset$weekofyear
modelTime = data.frame(time=seq(min(rsvTime) - 1 + eps, max(rsvTime), eps))

rsvCountyGroups = data.frame(county=rsvCounties) %>%
  mutate(
    group.all=TRUE,
    group.lowIncidence=county %in% c("Tolland", "Windham", "Middlesex", "Litchfield")
  )

ppxDuration = 24 # weeks

evalStrategy = function(start, end, time) {
  as.numeric((time > rsvWeek(start)) && (time < rsvWeek(end)))
}

evalOnsetStrategy = function(start, time) {
  as.numeric((time > rsvWeek(start)) && (time < rsvWeek(start + ppxDuration)))
}

evalOffsetStrategy = function(end, time) {
  as.numeric((time > rsvWeek(end - ppxDuration)) && (time < rsvWeek(end)))
}

ppxFixedStrategies = function(c=NULL) {
  stateOnset = stateThresholds$onset.median
  stateOffset = stateThresholds$offset.median

  if (!is.null(c)) {
    countyOnset = (thresholdsByCounty %>% filter(county==c))$onset.median
    countyOffset = (thresholdsByCounty %>% filter(county==c))$offset.median
  }

  list(
    # never=function(time) { 0 },
    # always=function(time) { 1 },
    aap=function(time) {
      evalOnsetStrategy(strptime("11-15", format="%m-%d")$yday / 7, time)
    },
    stateOnset=function(time) {
      evalOnsetStrategy(stateOnset, time)
    },
    stateOffset=function(time) {
      evalOffsetStrategy(stateOffset, time)
    },
    countyOnset=function(time) {
      evalOnsetStrategy(countyOnset, time)
    },
    countyOffset=function(time) {
      evalOffsetStrategy(countyOffset, time)
    }
  )
}

ppxSlidingStrategies = function(c=NULL, y=NULL) {
  if (!is.null(y) && !is.null(c)) {
    stateSlidingOnset = (stateThresholdsByWindow %>% filter(year==y))$onset.median
    stateSlidingOffset = (stateThresholdsByWindow %>% filter(year==y))$offset.median
  }

@@

```{r}
stateObs = dataset %>% 
  rename(time=weekofyear) %>% 
  group_by(time) %>% 
  summarize(rsv=sum(rsv)) %>% 
  mutate(rsv.cum=cumsum(rsv), rsv.cum.frac=rsv.cum / sum(rsv)) %>% 
  as.data.frame()

stateModel = gam(rsv ~ s(time, k=20, bs="cp", m=3), family=poisson, data=stateObs)

statePred = stateModel %>% 
  predict(type="response", newdata=modelTime, se.fit=TRUE)

statePred = modelTime %>% 
  cbind(data.frame(rsv.fit=statePred$fit, rsv.fit.se=statePred$se.fit)) %>%
  cbind(stateModel %>% outbreak.predict.timeseries(modelTime, function(model, params, newdata) { outbreak.calc.cum(model, params, newdata) })) %>%
  rename(rsv.cum.fit.lower=lower, rsv.cum.fit.upper=upper, rsv.cum.fit=median)

stateThresholds = stateModel %>%
  outbreak.predict.scalars(
    modelTime,
    function(model, params, newdata) { 
      outbreak.calc.thresholds(model, params, newdata$time, onset=seasonThreshold, offset=1 - seasonThreshold)
    }
  )

summarizeCountyGroups = function(data) {
  rbind(
      as.data.frame(data) %>%
        select(county, time, rsv), 
      as.data.frame(data) %>%
        filter(group.all==TRUE) %>%
        summarize(county="all", time=max(time), rsv=sum(rsv)),
      as.data.frame(data) %>%
        filter(group.lowIncidence==TRUE) %>%
        summarize(county="lowIncidence", time=max(time), rsv=sum(rsv))
    )  
}
```

@@

```{r}
obsByCounty = dataset %>% 
  rename(time=weekofyear) %>% 
  group_by(county, time) %>% 
  summarize(rsv=sum(rsv)) %>% 
  left_join(rsvCountyGroups, by="county") %>%
  group_by(time) %>%
  do(summarizeCountyGroups(.)) %>%
  group_by(county) %>% 
  mutate(rsv.cum=cumsum(rsv), rsv.cum.frac=rsv.cum / sum(rsv)) %>% 
  as.data.frame()

predByCounty = data.frame()
thresholdsByCounty = data.frame()
fixedStratUnprotectedByCounty = data.frame()

progress = txtProgressBar(max=length(levels(obsByCounty$county)), style=3)
idx = 0

for (c in levels(obsByCounty$county)) {
  countyObs = obsByCounty %>% filter(county==c)
  countyModel = gam(rsv ~ s(time, k=20, bs="cp", m=3), family=poisson, data=countyObs)

  countyPred = countyModel %>% 
    predict(type="response", newdata=modelTime, se.fit=TRUE)

  countyPred = modelTime %>% 
    cbind(data.frame(rsv.fit=countyPred$fit, rsv.fit.se=countyPred$se.fit)) %>%
    cbind(countyModel %>% outbreak.predict.timeseries(modelTime, function(model, params, newdata) { outbreak.calc.cum(model, params, newdata) })) %>%
    rename(rsv.cum.fit.lower=lower, rsv.cum.fit.upper=upper, rsv.cum.fit=median) %>%
    mutate(county=factor(c, levels=levels(obsByCounty$county)))

  predByCounty = predByCounty %>% rbind(countyPred)

  countyThresholds = countyModel %>%
    outbreak.predict.scalars(
      modelTime,
      function(model, params, newdata) { 
        outbreak.calc.thresholds(model, params, newdata$time, onset=seasonThreshold, offset=1 - seasonThreshold)
      }
    ) %>%
    mutate(county=factor(c, levels=levels(obsByCounty$county)))

  thresholdsByCounty = thresholdsByCounty %>% rbind(countyThresholds)

  countyUnprotected = countyModel %>%
    outbreak.predict.scalars(
      modelTime,
      function(model, params, newdata) { 
        outbreak.calc.unprotected(model, params, newdata$time, ppxFixedStrategies(c))
      }
    ) %>%
    mutate(county=factor(c, levels=levels(obsByCounty$county)))

  fixedStratUnprotectedByCounty = fixedStratUnprotectedByCounty %>% rbind(countyUnprotected)

  idx = idx + 1
  setTxtProgressBar(progress, idx)
}

>>>>

source("./RSV.R")
```

@@

```{r}
ppxSlidingStrategies = function(c=NULL, y=NULL) {
  if (!is.null(y) && !is.null(c)) {
    stateSlidingOnset = (stateThresholdsByWindow %>% filter(year==y))$onset.median
    stateSlidingOffset = (stateThresholdsByWindow %>% filter(year==y))$offset.median
  }

@@

@@
